FM project

### 2023-05-17 report.Rmd
updated

### 1.Introduction

@@

### 2. Time Series Graphics

@@

### 3. Time series decomposition

@@

### 5. Exponential smoothing.

@@

####5.1 Considering best ETS model

```{r setup, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
knitr::opts_chunk$set(echo = TRUE)

# Combine the transformed time series
combined_data <- merge(crude_oil_bx, nat_gas_bx, by = "Date", all = TRUE)
#plotting time series

#Determining the ETS Model for Crude OIL
ggplot(combined_data, aes(x = Date)) +
  geom_line(aes(y = Total_Exports.x, color = "Crude Oil")) +
  geom_line(aes(y = Total_Exports.y, color = "Natural Gas")) +
  ylab("Amount (Transformed)") +
  ggtitle("Crude Oil and Natural Gas Exports") +
  scale_color_manual(values = c("Crude Oil" = "red", "Natural Gas" = "green"))

@@

```{r cars}
#determining best ETS model
fit<- crude_oil_bx %>% 
  model(
    additive=ETS(Total_Exports~error("A")+trend("A")+season("A")),
    multiplicative = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
    random1 = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
  )

@@

addi_FC %>% 
  autoplot(
    crude_oil_agg) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

@@

5.2 Applying ETS() function

```{applying ETS formula}
#ETS model selection by R
crude_fit <- crude_oil_agg %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy
report(crude_fit)
# Print the selected ETS model
print(crude_fit$auto_crude_ets)
auto_crud%>% 
  autoplot(
    crude_oil_agg) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

```{bash}
#Plotting Residual
gg_tsresiduals(add)
gg_tsresiduals(crude_fit)
```

@@

>>>>

## 1.Introduction

@@

## 2. Time Series Graphics

@@

## 3. Time series decomposition

@@

## 5. Exponential smoothing.

@@

###5.1 Considering best ETS model

```{r setup, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

# Combine the transformed time series
combined_data <- merge(crude_oil_bx, nat_gas_bx, by = "Date", all = TRUE)

#plotting time series

#Determining the ETS Model for Crude OIL
ggplot(combined_data, aes(x = Date)) +
  geom_line(aes(y = Total_Exports.x, color = "Crude Oil")) +
  geom_line(aes(y = Total_Exports.y, color = "Natural Gas")) +
  ylab("Amount (Transformed)") +
  ggtitle("Crude Oil and Natural Gas Exports") +
  scale_color_manual(values = c("Crude Oil" = "red", "Natural Gas" = "green"))
```

@@

```{r cars, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}}
#determining best ETS model
fit <- crude_oil_bx %>% 
  model(
    additive_ETS=ETS(Total_Exports~error("A")+trend("A")+season("A")),
    multiplicative_ETS = ETS(Total_Exports ~ error("M") + trend("M") + season("M")),
    random1_ETS = ETS(Total_Exports ~ error("M") + trend("A") + season("A")),
  )

@@

addi_fc<- fit %>% select(additive_ETS) %>% forecast(h=36)

addi_fc %>% 
  autoplot(
    crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

@@

### 5.2 Applying ETS() function

```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#ETS model selection by R
crude_fit <- crude_oil_bx %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy
report(crude_fit)
# Print the selected ETS model
print(crude_fit$auto_crude_ets)

crude_fit$auto_crude_ets %>% 
  autoplot(crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

```{r bash, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#Plotting Residual
gg_tsresiduals(addi_fc)
gg_tsresiduals(crude_fit)
```

@@

### 6.1. Box-Jenkins methodology

### Random walk

The Box-Cox transformation was already applied to stabilize the variance in chapter [3.2. Transformations]. Other than that no transformations were applied on neither of time series.

Following a definition of stanioarity provided by Rob J Hyndman and George Athanasopoulos in Forecasting: Principles and Practice (3rd ed) **("A stationary time series is one whose statistical properties do not depend on the time at which the series is observed")** it is obvious that our time series is non stationary since it has clear trend and some seasonality.

We start by performing Unit Root test to identify which methods are suitable in our case
```{r, Unit root test}

crude_oil_bx %>% ACF(Total_Exports) %>% 
  autoplot() + labs(title = "ACF after first order difference")

# KPSS test

crude_oil_bx  %>% 
  features(Total_Exports, unitroot_kpss)

#Testing for seasonal differencing

crude_oil_bx %>% 
  features(Total_Exports, unitroot_nsdiffs)

```
We see that data, as expected, needs differencing but no seasonal differencing will be applied since unitroot_nsdiffs test showed that there is no need of that.

```{r, first order difference}

crude_oil_bx <- crude_oil_bx %>% 
  mutate(diff_total_exports = difference(Total_Exports))

crude_oil_bx  %>% 
  features(diff_total_exports, unitroot_kpss)

```
After conducting first order difference we see that data is now stationary.

```{r}

crude_oil_bx %>% ACF(diff_total_exports) %>% 
  autoplot() + labs(title = "ACF of stationar crude oil exports")

crude_oil_bx %>% PACF(diff_total_exports) %>% 
  autoplot() + labs(title = "PACF of stationar crude oil exports ")
```
